Cybersecurity impacts at the beginning\\ of Covid-19 pandemic in Italy

### 2021-03-06 article.tex
added istat reference

OAD 2020 survey\footnote{The report is written in Italian, only the Executive Summary is in English: 186 A4 pages, 148 images and graphics, 11 Chapters
(147 A4 pages) and 9 Attachments (39 A4 pages)}~\cite{oad20} took place during the Covid-19 pandemic, and it covered the entire 2019 and the first quarter 
of 2020, when in Italy the pandemic exploded. The data comparison of 2019 and the first quarter 2020, provided by the same respondents pool, has a statistical relevance and provides a 
clear indication of the Covid-19 pandemic on the intentional cyber attacks in Italy. This pandemic has been the trigger for a wide range of cyber attacks, mainly 
caused by the sudden - and in part unprepared - passage of many to agile working from home and to a strong use of every type of IT service on Internet, mainly
derived by the mobility lockdowns imposed by the Italian Authorities. The emerged respondents’ pool covers all the product sectors, including Public Administrations,
even if the majority of the respondents’ companies belong to the ICT sector (30.8\%). The 2020 pool is well balanced for the size of the organizations, in terms
of employees, between those below 250 and the largest ones. It must be taken into account that in Italy 99.91\% \cite{istat20} %TODO: manca fonte/citazione
of enterprises are SMEs, under 250 employees, and 95\% are under ten employees. The OAD 2020 survey is therefore able to consider small and 
tiny organizations, which are thee vast majority in Italy, and which are not normally considered in other cyber security surveys: 57.5\% of the OAD 2020 respondents
belong to structures with less than 250 employees, and of these 22.1\% under 10.

>>>>

OAD 2020 survey\footnote{The report is written in Italian, only the Executive Summary is in English: 186 A4 pages, 148 images and graphics, 11 Chapters
(147 A4 pages) and 9 Attachments (39 A4 pages)}~\cite{oad20} took place during the Covid-19 pandemic, and it covered the entire 2019 and the first quarter 
of 2020, when in Italy the pandemic exploded. The data comparison of 2019 and the first quarter 2020, provided by the same respondents pool, has a statistical relevance and provides a 
clear indication of the Covid-19 pandemic on the intentional cyber attacks in Italy. This pandemic has been the trigger for a wide range of cyber attacks, mainly 
caused by the sudden - and in part unprepared - passage of many to agile working from home and to a strong use of every type of IT service on Internet, mainly
derived by the mobility lockdowns imposed by the Italian Authorities. The emerged respondents’ pool covers all the product sectors, including Public Administrations,
even if the majority of the respondents’ companies belong to the ICT sector (30.8\%). The 2020 pool is well balanced for the size of the organizations, in terms
of employees, between those below 250 and the largest ones. It must be taken into account that in Italy 99.91\% \cite{istat21}
of enterprises are SMEs, under 250 employees, and 95\% are under ten employees. The OAD 2020 survey is therefore able to consider small and 
tiny organizations, which are thee vast majority in Italy, and which are not normally considered in other cyber security surveys: 57.5\% of the OAD 2020 respondents
belong to structures with less than 250 employees, and of these 22.1\% under 10.
